---
title: "Retrieve standard data from Seatrack"
author: "Jens Åström"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    highlight: tango
  rmarkdown::html_vignette:
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{R.rsp::asis}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy = T)
require(tidyverse)
```

Connecting to the database
=======
Users at NINA and the Polarinstitute (using a computer that is within these networks' IP-addresses) can connect to the Seatrack database. It is a PostgreSQL (9.6) database answering to the address `seatrack.nina.no`, on the standard port 5432. Users should use their individual login user names and passwords. Contact Jens Åström (jens.astrom@nina.no) for details about usernames and passwords.

This instruction deals with the preferred way of connecting to the database, using R and the `seatrackR` package. Another option is to connect through a dedicated database management software, such as Pgadmin3 (or 4), HeidiSQL, or similar. Some users may prefer to use the MS Access interface.


To simplify the connection, use the convenience function `connectSeatrack`. This creates a connection named `con` by using the packages `DBI` and `RPostgres`.


```{r}
require(seatrackR)

connectSeatrack(Username =  "testreader", Password = "testreader")
```


Custom queries
==============
As of now, 4 functions exist to retreive data from the database through prebuilt queries. Apart from that, users are free to use their own queries through the functions in `DBI` and `dplyr`, using the connection named `con` made by the `connectSeatrack()`-function.

It is perfecty fine to download data through your own custom queries. Creating interesting queries requires some knowledge about the structure of the database however. Pgadmin3(4) would be a useful tool to get further info on that. For now, we show a simple query involving just one table. Here we get the different locations currently recorded from the Faroe Islands (Coordinates not updated).  Note that you have to load the `DBI` package and use its query functions.

```{r}
require(DBI)

myQuery <- "SELECT * from metadata.location
              WHERE colony_int_name = 'Faroe Islands'"

faroeLocations <- dbGetQuery(con, myQuery)
head(faroeLocations)

```


Position data
===============
The primary data of the positions of the birds is stored in the table `positions.postable`. This includes all entered positions in the database. 

The `getPosdata` function retrieves this table, with options to subselect only specific species, colonies, responsible contact person, specific ring numbers, and years. There is also an option to limit the records to a set number of rows, and to load the position coordinates as a spatial object.

```{r}

eynhallowPositions <- getPosdata(selectColony = "Eynhallow",
                                 loadGeometries = T)
eynhallowPositions
```


```{r}
plot(eynhallowPositions["ring_number"])
```


Position data for export
==================
The data sent to the Polar institute also have the subspecies names added to the records. The export ready positions data can be retrieved most easily through a specific export view. Note that this will download all records, and will take some time. Note that this export does not contain information on the used and deleted uuids. 

```{r, eval = F}
newExport <- dbReadTable(con, Id(schema = "views", table = "export"))
nrow(newExport)
write.csv(newExport, file = "seatrack_export_2018-08-09.csv")
```

If you are interested in knowing separate old, deleted rows, these are found in the table `positions.deleted_uuid`.

```{r, eval = F}
deletedUuids <- dbReadTable(con, Id(schema = "positions", table =  "deleted_uuid"))
nrow(deletedUuids)
write.csv(deletedUuids, file = "deletedUuids_2018-08-09.csv")

```


Other functions for download
================
There are some more convenience functions for retrieving information from the database as well. Here follows a quich demo. 

The `getFileArchiveSummary` function
--------------
This function pulls together data from several tables with focus on the file archive. It should contain enough information to know what the individual raw files contain.

```{r}
eynhallowFiles <- getFileArchiveSummary(selectColony = "Eynhallow")
eynhallowFiles
```


The `getIndividInfo` function
-------------
This function summarizes all observation data for the individual birds. We can subselect the colony and year interval the bird where tracked.

```{r}
hornoyaIndivids <- getIndividInfo(selectColony = "Hornoya",
                                    selectYear = "2014_15")
hornoyaIndivids
```

!Note the weird duplicate records here! **TO BE FIXED**

```{r}
hornoyaIndivids %>% print(width = Inf)
```


Commonly used info
-------------
I have made a couple of views for som common information, that are displayed on the shiny app http://view.nina.no/seatrack/. These can be found like this as well.

```{r}
shorttable <- dbReadTable(con, Id(schema = "views", table = "shorttable"))
shorttable
```
```{r}
shorttableeqfilter3 <- dbReadTable(con, Id(schema = "views", table = "shorttableeqfilter3"))
shorttableeqfilter3
```




