---
output: 
  html_document:
      toc: true
      toc_depth: 3
      toc_float: true
params:
  request_name: "myrequest2025"
  file_list: !r c("locations.parquet")
  species: !r c("seagull")
  colonies: !r c("big cliff by the sea")
  times: !r as.Date(c("2020-01-01","2020-02-02"))
  data_types: !r c("location")
  creation_date: !r Sys.Date()
  data_dir: ""
  notes: ""
  additional_files: !r list()
title: "`r paste(params$request_name, 'README', params$creation_date)`"
---

```{r, echo = FALSE, message = FALSE}
library(yaml)
library(knitr)

data_dir <- file.path(params$data_dir, "data") # to use in the rest of the examples

pos_path <- params$file_list[grepl("posdata", params$file_list)]
imm_path <- params$file_list[grepl("immersion", params$file_list)]
temp_path <- params$file_list[grepl("temperature", params$file_list)]
light_path <- params$file_list[grepl("light", params$file_list)]
indiv_path <- params$file_list[grepl("morph_and_breeding", params$file_list)]

md_list <- function(values){
  md_items <- sapply(values, function(value){
    knitr::knit_expand(text = "| - *{{value}}*")
    })
  return(paste(md_items, collapse = "/n"))
}

has_pos <- has_imm <- has_light <- has_temp <- has_ind <- TRUE
file_list <- c()

if (length(params$additional_files) > 0) {
  for (additional_file in params$additional_files){
    filename <- basename(additional_file$path)
    file_list <- c(file_list, '| -- {{additional_file$description}} (<span style="color: #ABB2B9;">{{filename}}</span>)')
  }
}
file_list <- c(file_list, "| -- data")

if (length(pos_path) > 0) {
  has_pos <- TRUE
  file_list <- c(file_list, '| &nbsp;&nbsp;&nbsp;&nbsp;|-- Positional data (<span style="color: #ABB2B9;">{{pos_path}}</span>)')
}
if (length(imm_path) > 0) {
  has_imm <- TRUE
  file_list <- c(file_list, '| &nbsp;&nbsp;&nbsp;&nbsp;|-- Standardized immersion data (<span style="color: #ABB2B9;">{{imm_path}}</span>)')
}
if (length(light_path) > 0) {
  has_light <- TRUE
  file_list <- c(file_list, '| &nbsp;&nbsp;&nbsp;&nbsp;|-- Light data (<span style="color: #ABB2B9;">{{light_path}}</span>)')
}
if (length(temp_path) > 0) {
  has_temp <- TRUE
  file_list <- c(file_list, '| &nbsp;&nbsp;&nbsp;&nbsp;|-- Temperature data (<span style="color: #ABB2B9;">{{temp_path}}</span>)')
}
if (length(indiv_path) > 0) {
  has_ind <- TRUE
  file_list <- c(file_list, '| &nbsp;&nbsp;&nbsp;&nbsp;|-- Individual information data (<span style="color: #ABB2B9;">{{indiv_path}}</span>)')
}



file_list <- paste(knitr::knit_expand(text = file_list), collapse = "\n")


species_list <- md_list(params$species)
colony_list <- md_list(params$colonies)
time_span <- knitr::knit_expand(text = "| - *{{params$times[1]}}* - *{{params$times[2]}}*")
additional_notes <- knitr::knit_expand(text = "| {{params$notes}}")

# Load schema
schema_dict <- read_yaml(system.file("yaml", "output_schema.yaml", package = "seatrackR"))

render_schema <- function(dataset_key, dict) {
  dataset <- dict$datasets[[dataset_key]]
  title <- gsub("_", " ", tools::toTitleCase(dataset_key))
  header <- knitr::knit_expand(text = "### {{title}} \n\n {{dataset$description}}")

  bullets <- sapply(dataset$columns, function(col) {
      knitr::knit_expand(text = "- **{{col$name}}** *(type:{{col$type}})* \n  {{col$description}} ")
    })
  all_desc <- paste(c(header, bullets), collapse = "\n\n")
  return(all_desc)
}

schema_md <- paste(sapply(names(schema_dict$datasets), render_schema, dict = schema_dict), collapse = "\n\n")
```
## Description
| ___________________________________________________________________________________________________
| This readme contains an overview of the files contained within this data package and instructions on how to use them in R. 
`r knitr::knit(text = additional_notes)`
|
| *For more information about the project, please visit [SEATRACK's website](http://www.seapop.no/en/seatrack/)*
| *contact re. General data queries: julian.evans@nina.no*
|
| *For more information about the processing of GLS data, please see this [report](https://brage.nina.no/nina-xmlui/handle/11250/2735757)*
| *contact re. GLS data processing: vegard.brathen@nina.no*
|
| ___________________________________________________________________________________________________
## Package overview
| ___________________________________________________________________________________________________
| **The dataset consists of:**
| **Files**:
| -- This readme (<span style="color: #ABB2B9;">`r paste0("README_", params$request_name, "_", params$creation_date, ".html")`</span>)
`r knitr::knit(text = file_list)`
|
| **Time range**
`r knitr::knit(text = time_span)`
|
| **Species**: 
`r knitr::knit(text = species_list)`
| 
| **Colonies**: 
`r knitr::knit(text = colony_list)`
| ___________________________________________________________________________________________________


## Examples of use

### Prepare workspace & load data

```{r echo=TRUE, eval = FALSE, message=FALSE}
# Load libraries
library(arrow) # for loading parquet files
library(dplyr) # for easily working with data
library(lubridate) # for some enhanced date functionality
library(ggplot2) # for plotting
library(sf) # for GIS work
library(maps)

# deactivate spherical geometry (s2) for sf
sf::sf_use_s2(FALSE)

# point R to the folder where the data is, e.g.
data_dir <- "path/where/you/unzipped/files/data"
```

```{r echo=FALSE, message=FALSE}
# THE ABOVE BLOCK IS FAKE TO SHOW HOW TO SET THE 
# Load libraries
library(arrow) # for loading parquet files
library(dplyr) # for easily working with data
library(lubridate) # for some enhanced date functionality
library(ggplot2) # for plotting
library(sf) # for GIS work
library(maps)

# deactivate spherical geometry (s2) for sf
sf::sf_use_s2(FALSE)
```

The arrow package can be used to load data files

```{r generate_load_files, echo = FALSE}

# MODIFY THIS TO ONLY READ IN THE FILES PRESENT
load_file_chunk <- knitr::knit_expand(text = c(
  "```{r load_files}",
  'mylocs <- read_parquet(file.path(data_dir, "{{pos_path}}"))',
  'immersion <- read_parquet(file.path(data_dir, "{{imm_path}}"))',
  'light <- read_parquet(file.path(data_dir, "{{light_path}}"))',
  'individs <- read_parquet(file.path(data_dir, "{{indiv_path}}"))',
  "```"
))

```

`r knitr::knit(text = load_file_chunk)`

### Explore data

Some examples on how to quickly explore your data.

```{r echo=TRUE, message=TRUE}
mylocs %>%
  dplyr::group_by(colony, year_tracked, age_deployment) %>%
  dplyr::summarise(n_birds_tracked = n_distinct(individ_id))
```

```{r echo = TRUE, message = TRUE}
immersion %>%
  dplyr::mutate(year = year(date_time)) %>%
  dplyr::group_by(year) %>%
  dplyr::summarise(n_birds_tracked = n_distinct(individ_id))
```

```{r echo = TRUE, message = TRUE}
light %>%
  dplyr::mutate(year = year(date_time)) %>%
  dplyr::group_by(year) %>%
  dplyr::summarise(n_birds_tracked = n_distinct(individ_id))
```

```{r echo=TRUE, message=TRUE}
individs %>%
  dplyr::group_by(colony, year_tracked) %>%
  dplyr::summarise(n_birds_tracked = n_distinct(individ_id))
```
### Map positions

Some examples on how to quickly map your data.

```{r echo = TRUE, message = TRUE}
mylocs %>%
  dplyr::mutate(year = year(date_time)) %>%
  sf::st_as_sf(coords = c("lon", "lat"), crs = 4326) ->
mylocs

worldmap <- ggplot2::map_data("world")

ggplot() +
  geom_polygon(data = worldmap, aes(x = long, y = lat, group = group)) +
  geom_sf(data = mylocs[mylocs$eqfilter == 1, ], aes(colour = age_deployment), size = .1) +
  coord_sf(
   lims_method = "geometry_bbox"
  ) +
  theme_light() +
  theme(legend.position = "bottom")
```


### Plot immersion data

An example of plotting immersion data.
```{r echo = TRUE, message = TRUE}
one_individ <- immersion[immersion$individ_id %in% unique(immersion$individ_id)[1], ]
one_year <- one_individ[year(one_individ$date_time) %in% unique(year(one_individ$date_time))[2], ]
one_year$doy <- strftime(one_year$date_time, format = "%j")
one_year <- aggregate(std_conductivity ~ doy, one_year, mean)
plot(one_year$doy, one_year$std_conductivity,
  type = "l", col = "lightblue", lwd = 2,
  xlab = "day of the year", ylab = "mean daily immersion (0-1)", main = unique(immersion$individ_id)[1]
)
``` 

### Plot light data

An example of plotting light data.
```{r echo = TRUE, message = TRUE}
one_individ <- light[light$individ_id %in% unique(light$individ_id)[1], ]
one_year <- one_individ[year(one_individ$date_time) %in% unique(year(one_individ$date_time))[], ]
one_year$doy <- strftime(one_year$date_time, format = "%j")
one_year <- aggregate(raw_light ~ doy, one_year, sum)
plot(one_year$doy, one_year$raw_light, type = "l", col = "lightblue", lwd = 2, xlab = "day of the year", ylab = "sum daily light", main = unique(light$individ_id)[1])
``` 

## Detailed file description

`r knitr::knit(text = schema_md)`

