---
output: 
  html_document:
      toc: true
      toc_depth: 3
      toc_float: true
params:
  request_name: "myrequest2025"
  file_list: !r c("locations.parquet")
  species: !r c("seagull")
  colonies: !r c("big cliff by the sea")
  times: !r as.Date(c("2020-01-01","2020-02-02"))
  data_types: !r c("location")
  creation_date: !r Sys.Date()
  data_dir: ""
  notes: ""
  additional_files: !r list()
title: "`r paste(params$request_name, 'README', params$creation_date)`"
---

```{r, setup , echo = FALSE, message = FALSE, warning = FALSE}
data_dir <- file.path(params$data_dir, "data") # to use in the rest of the examples

# Detect paths
pos_path <- params$file_list[grepl("GLS_positional_data", params$file_list)]
imm_path <- params$file_list[grepl("immersion", params$file_list)]
temp_path <- params$file_list[grepl("temperature", params$file_list)]
light_path <- params$file_list[grepl("light", params$file_list)]
indiv_path <- params$file_list[grepl("individual_data", params$file_list)]
pop_map_paths <- params$file_list[grepl("Abundance_Model", params$file_list)]

colonies <- params$colonies

md_list <- function(values) {
  if(is.null(values)){
    return("")
  }
  md_items <- sapply(values, function(value) {
    knitr::knit_expand(text = "| - *{{value}}*")
  })
  return(paste(md_items, collapse = "\n"))
}

has_pos <- has_imm <- has_light <- has_temp <- has_ind <- has_pop <- FALSE
file_list <- c()

if (length(params$additional_files) > 0) {
  for (additional_file in params$additional_files) {
    filename <- basename(additional_file$path)
    file_list <- c(file_list, '| -- {{additional_file$description}} (<span style="color: #ABB2B9;">{{filename}}</span>)')
  }
}
file_list <- c(file_list, "| -- data")

included_data <- c()
if (length(pos_path) > 0) {
  has_pos <- TRUE
  included_data <- c(included_data, "GLS_positional_data")
  file_list <- c(file_list, '| &nbsp;&nbsp;&nbsp;&nbsp;|-- Positional data (<span style="color: #ABB2B9;">{{pos_path}}</span>)')
}
if (length(imm_path) > 0) {
  has_imm <- TRUE
  included_data <- c(included_data, "immersion_data")
  file_list <- c(file_list, '| &nbsp;&nbsp;&nbsp;&nbsp;|-- Standardized immersion data (<span style="color: #ABB2B9;">{{imm_path}}</span>)')
}
if (length(light_path) > 0) {
  has_light <- TRUE
  included_data <- c(included_data, "light_data")
  file_list <- c(file_list, '| &nbsp;&nbsp;&nbsp;&nbsp;|-- Light data (<span style="color: #ABB2B9;">{{light_path}}</span>)')
}
if (length(temp_path) > 0) {
  has_temp <- TRUE
  included_data <- c(included_data, "temperature_data")
  file_list <- c(file_list, '| &nbsp;&nbsp;&nbsp;&nbsp;|-- Temperature data (<span style="color: #ABB2B9;">{{temp_path}}</span>)')
}
if (length(indiv_path) > 0) {
  has_ind <- TRUE
  included_data <- c(included_data, "individual_data")
  file_list <- c(file_list, '| &nbsp;&nbsp;&nbsp;&nbsp;|-- Individual information data (<span style="color: #ABB2B9;">{{indiv_path}}</span>)')
}
if (length(pop_map_paths) > 0) {


  species_names <- data.frame(latin = c(
  "Alle alle", 
  "Fratercula arctica", 
  "Fulmarus glacialis", 
  "Rissa tridactyla", 
  "Uria aalge", 
  "Uria lomvia"), 
  common = c(
    "Little auk",
   "Atlantic puffin", 
   "Northern fulmar", 
   "Black-legged kittiwake", 
   "Common guillemot",
   "Brünnich's guillemot"))
  included_data <- c(included_data, "population_maps")
  if(is.null(colonies)){
    colonies <- c()
  }
  has_pop <- TRUE
  for (pop_map_path in pop_map_paths) {
    file_name <- basename(pop_map_path)
    split_file_name <- strsplit(file_name, "_")[[1]]
    latin <- paste(split_file_name[c(4, 5)], collapse = " ")
    common <- species_names$common[species_names$latin == latin]
    version <- paste(split_file_name[c(7, 8)], collapse = ".")
    popmap_text <- '| &nbsp;&nbsp;&nbsp;&nbsp;|-- SEATRACK population map {{version}} - {{common}} (<span style="color: #ABB2B9;">{{pop_map_path}}</span>)'
    file_list <- c(file_list, knitr::knit_expand(text = popmap_text))
    loaded_netcdf <- nc_open(file.path(data_dir, pop_map_path))
    colony_names <- ncvar_get(loaded_netcdf, "SmcolName")
    colonies <- c(colonies, colony_names)
  }
}



file_list <- paste(knitr::knit_expand(text = file_list), collapse = "\n")

species_list <- md_list(params$species)
colony_list <- md_list(unique(colonies))
time_span <- knitr::knit_expand(text = "| - *{{params$times[1]}}* - *{{params$times[2]}}*")
additional_notes <- knitr::knit_expand(text = "| {{params$notes}}")

# Load schema
schema_dict <- yaml::read_yaml(system.file("yaml", "output_schema.yaml", package = "seatrackR"))

render_schema <- function(dataset_key, dict) {
  dataset <- dict$datasets[[dataset_key]]
  title <- gsub("_", " ", tools::toTitleCase(dataset_key))
  header <- knitr::knit_expand(text = "### {{title}} \n\n {{dataset$description}}")

  bullets <- sapply(dataset$columns, function(col) {
    knitr::knit_expand(text = "- **{{col$name}}** *(type: {{col$type}})* \n  {{col$description}} ")
  })
  all_desc <- paste(c(header, bullets), collapse = "\n\n")
  return(all_desc)
}

included_schema <- names(schema_dict$datasets)[names(schema_dict$datasets) %in% included_data]

schema_md <- paste(sapply(included_schema, render_schema, dict = schema_dict), collapse = "\n\n")

image_path <- system.file("img", "SEATRACK logo landscape.jpg", package = "seatrackR")


```

```{r include-package-image, echo=FALSE}
    knitr::include_graphics(image_path)
```

## Description
| ___________________________________________________________________________________________________
| This readme contains an overview of the files contained within this data package and instructions on how to use them in R. 
`r knitr::knit(text = additional_notes)`
|
| *For more information about the project, please visit [SEATRACK's website](http://www.seapop.no/en/seatrack/)*
| *contact re. General data queries: julian.evans@nina.no*
|
`r if(has_pos) {"| *For more information about the processing of GLS data, please see this [report](https://brage.nina.no/nina-xmlui/handle/11250/2735757)*"}`
`r if(has_pos) {"| *contact re. GLS data processing: vegard.brathen@nina.no*"}`
|
`r if(has_pop) {"| *For more information about the creation of population maps, please see this [paper](https://www.int-res.com/articles/meps_oa/m676p255.pdf)*"}`
`r if(has_pop) {"| *contact re. Population maps/IRMA: arnaud.tarroux@nina.no*"}`
| ___________________________________________________________________________________________________
## Package overview
| ___________________________________________________________________________________________________
| **The dataset consists of:**
| **Files**:
| -- This readme (<span style="color: #ABB2B9;">`r paste0("README_", params$request_name, "_", params$creation_date, ".html")`</span>)
`r knitr::knit(text = file_list)`
|
| **Time range**
`r knitr::knit(text = time_span)`
|
| **Species**: 
`r knitr::knit(text = species_list)`
| 
| **Colonies**: 
`r knitr::knit(text = colony_list)`
| ___________________________________________________________________________________________________


## Examples of use

```{r, show_examples, echo=FALSE, results='asis'}
if (has_pos) {
  res <- knitr::knit_child(system.file("Rmd", "EXAMPLES_locations.Rmd", package = "seatrackR"),
    quiet = TRUE, envir = environment()
  )
  cat(res, sep = "\n")
}

if (has_ind) {
  res <- knitr::knit_child(system.file("Rmd", "EXAMPLES_individuals.Rmd", package = "seatrackR"),
    quiet = TRUE, envir = environment()
  )
  cat(res, sep = "\n")
}

if (has_light) {
  res <- knitr::knit_child(system.file("Rmd", "EXAMPLES_light.Rmd", package = "seatrackR"),
    quiet = TRUE, envir = environment()
  )
  cat(res, sep = "\n")
}

if (has_imm) {
  res <- knitr::knit_child(system.file("Rmd", "EXAMPLES_immersion.Rmd", package = "seatrackR"),
    quiet = TRUE, envir = environment()
  )
  cat(res, sep = "\n")
}

if (has_pop) {
  res <- knitr::knit_child(system.file("Rmd", "EXAMPLES_popmap.Rmd", package = "seatrackR"),
    quiet = TRUE, envir = environment()
  )
  cat(res, sep = "\n")
}
```

## Detailed file description

`r knitr::knit(text = schema_md)`

